# Class Activity 15

```{r}
# load the necessary libraries
library(tidyverse)
library(dplyr)
library(stringr)


energy <- read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/energy.csv",
                    col_type = cols(
                     .default = col_double(), 
                      Timestamp = col_datetime(format = ""),
                      dayWeek = col_factor(levels=c("Mon","Tues","Wed","Thurs","Fri","Sat","Sun"))
                     ))
```


## Group Activity 1

### a. `if` and `for` loop

Write a for loop to iterate over the columns of the 'energy' dataset and print the names of all columns containing the string "House". Please use the function `colnames()` to extract the column names and store the results in a list.


<details>
<summary class="answer">Click for answer</summary>
*Answer:*

```{r}
# Create an empty list to store the column names
house_columns <- list()

# Iterate over the columns of the 'energy' dataset
for (i in seq_along(colnames(energy))) {
  col_name <- colnames(energy)[i]
  
  # Check if the column name contains the string "House"
  if (str_detect(col_name, "House")) {
    # Add the column name to the list
    house_columns[[length(house_columns) + 1]] <- col_name
  }
}

# Print the list of house columns
house_columns <- unlist(house_columns)
house_columns
```

</details>

### b. `for` loop and mean

Using a for loop, calculate and print the mean energy consumption of houses you identified in part a.


<details>
<summary class="answer">Click for answer</summary>
*Answer:*

```{r}
# Assuming the house_columns vector from the previous step

# Create an empty numeric vector to store the mean energy consumption
mean_energy_consumption <- numeric()

# Iterate over the house_columns vector
for (house_col in house_columns) {
  # Calculate the mean energy consumption for the current house column
  mean_val <- mean(energy[[house_col]], na.rm = TRUE)
  
  # Add the mean energy consumption to the vector
  mean_energy_consumption <- c(mean_energy_consumption, mean_val)
}

# Combine the house names and mean energy consumption into a dataframe
house_mean_energy <- bind_cols(House = house_columns, MeanEnergyConsumption = mean_energy_consumption)

# Print the dataframe
house_mean_energy %>%  knitr::kable()
```

</details>

------------------------------------------

## Group Activity 2


**1.** Make a data frame of quantiles for `energy` buildings in columns 9-90 (you will need `na.rm = TRUE`)


<details>
<summary class="answer">Click for answer</summary>
*Answer:*

```{r}
qdf <- energy %>% select(9:90) %>%
  map_dfc(quantile, probs = seq(.1,.9,.1), na.rm = TRUE)
qdf
```

</details>

**2.**  Add a variable to identify the quantile


<details>
<summary class="answer">Click for answer</summary>
*Answer:*

```{r}
qdf <- energy %>% select(9:90) %>%
  map_dfc(quantile, probs = seq(.1,.9,.1), na.rm = TRUE) %>%  
  mutate(stat = str_c("quantile_", seq(10,90,10))) 
qdf
```

</details>

**3.**  Reshape the data frame to make variables `stat` (describing the quantile), `building` and `quant` (quantile value)


<details>
<summary class="answer">Click for answer</summary>
*Answer:*


```{r}
qdf <- energy %>% select(9:90) %>%
  map_dfc(quantile, probs = seq(.1,.9,.1), na.rm = TRUE) %>%  
  mutate(stat = str_c("quantile_", seq(10,90,10))) %>%   
  pivot_longer(names_to = "building", values_to = "quantiles", 1:82)
qdf
```

</details>

**4.**  Plot the KWH value for each quantile on the x-axis for the buildings `Sayles-Hill`, `Language_&_Dining_Center`, `Olin_Hall_of_Science`

<!--

<details>
<summary class="answer">Click for answer</summary>
*Answer:*

```{r}
qdf %>% 
  filter(building %in% c("Sayles-Hill" ,"Language_&_Dining_Center", "Olin_Hall_of_Science"))  %>%  
  ggplot(aes(x=quantiles, y=parse_number(stat), color=building)) +
  geom_point() + 
  geom_line(aes(group=building)) + 
  labs(y="Percentile (%)",x="KWH") +
  scale_y_continuous(breaks=seq(10,90,by=10))
```


</details>

-->
