# Class Activity 12

```{r}
# load the necessary libraries
library(stringr)
library(dplyr)
library(readr)
```


In this tutorial, we will learn about string manipulations using regular expressions and the `stringr` library in R. We will cover different examples and use cases to help you understand the concepts and functions related to string manipulation.


## Group Activity 1

```{r}
x <- "My SSN is 593-29-9502 and my age is 55"
y <- "My phone number is 612-643-1539"
z <- "My old SSN number is 39532 9423."
out <- str_flatten(c(x,y,z), collapse = ". ")
```

### a. What characters in `x` will `str_view_all(x, "-..-")` find?

<details>
<summary class="answer">Click for answer</summary>
*answer:*

The pattern searches for a dash, followed by any two characters, followed by another dash. In x, it finds "-29-" which is a part of the SSN.

```{r}
str_view_all(x, "-..-")
```


</details>

### b. What pattern will `str_view_all(x, "-\\d{2}-")` find?

<details>
<summary class="answer">Click for answer</summary>
*answer:*

The pattern searches for a dash, followed by two digits, followed by another dash. In x, it finds the same "-29-" as in the previous example, which is a part of the SSN.


```{r}
str_view_all(x, "-\\d{2}-")  # "-" then 2 digits then "-"
```


</details>

### c. What pattern will `str_view_all(out, "\\d{2}\\.*")` find?

<details>
<summary class="answer">Click for answer</summary>
*answer:*

The pattern searches for two digits followed by an optional period. In out, it finds "55" and "55.", which represent the age in the first sentence.

```{r}
str_view_all(out, "\\s\\d{2}\\.")  # 2 digits then "."
```


</details>

### d. Use `str_view_all` to determine the correct regex pattern to identify all SSN in `out`

We can get the SSN with the usual format (###-##-####) with a regex that has 3, 2, and 4 digits separated by a dash. 

```{r}
str_view_all(out,"([0-8]\\d{2})-(\\d{2})-(\\d{4})")
```

This misses the oddly formatted SSN in the third entry. Rather than use a dash, we can specify the divider as `[-\\s]?` which allows either 0 or 1 occurrences of either a dash or space divider:

```{r }
str_view_all(out,"([0-8]\\d{2})[-\\s]?(\\d{2})[-\\s]?(\\d{4})")
```

<details>
<summary class="answer">Click for answer</summary>
*answer:*

The first pattern finds the SSNs in the standard format (###-##-####) by searching for 3 digits, a dash, 2 digits, another dash, and 4 digits. The second pattern does the same but allows for a space instead of a dash as a divider. It finds all SSNs in out, including the oddly formatted one in the third sentence.

</details>

### e. Write a regular expression to extract dates in the format YYYY-MM-DD from a given text.

```{r}
date_pattern <- "\\d{4}-\\d{2}-\\d{2}"
text <- "The event will take place on 2023-07-20 and end on 2023-07-22."
str_extract_all(text, date_pattern)
```


<details>
<summary class="answer">Click for answer</summary>
*Answer:* The pattern searches for 4 digits, a dash, 2 digits, another dash, and 2 digits. In the given text, it finds the dates "2023-07-20" and "2023-07-22".


</details>

### f. Write a regular expression to extract all words that start with a capital letter in a given text.


```{r}
capital_pattern <- "\\b[A-Z][a-zA-Z]*\\b"
text <- "Alice and Bob went to the Market to buy some Groceries."
str_extract_all(text, capital_pattern)
```

<details>
<summary class="answer">Click for answer</summary>
*Answer:* The pattern searches for a word boundary, followed by an uppercase letter, and then any sequence of letters. In the given text, it finds the words "Alice", "Bob", "Market", and "Groceries".

</details>

-----------------------------------------------------------------------------------

## Group Activity 2

### a. Let's deal with a number string that is longer than 9 digits.

```{r }
ssn <- "([0-8]\\d{2})[-\\s]?(\\d{2})[-\\s]?(\\d{4})"
test <- c("123-45-67890","1123 45 6789")
str_view_all(test, ssn)
```

This example captures a 9-digit string as an SSN, but these strings are longer than 9 digits and may not represent an SSN. One way to deal with this is to use the negative lookbehind `?<!` and negative lookahead `?!` operators to ensure that the identified 9-digit string does not have a leading 0 or does not contain more digits.

If we "look behind" from the start of the SSN, we should not see another digit:

```{r }
str_view_all(test, "(?<!\\d)([0-8]\\d{2})[-\\.\\s]?(\\d{2})[-\\.\\s]?(\\d{4})")
```

And if we "look ahead" from the end of the SSN, we should not see another digit:

```{r}
str_view_all(test, "(?<!\\d)([0-8]\\d{2})[-\\.\\s]?(\\d{2})[-\\.\\s]?(\\d{4})(?!\\d)")
```


For parts b and c, consider the following string. 

```{r}
string1 <- "100 dollars 100 pesos"
```


### b. Explain why the following matches the first 100 and not the second.

<details>
<summary class="answer">Click for answer</summary>
*answer:* It looks for one or more digits followed by a space and `dollars`


```{r}
str_view(string1, "\\d+(?= dollars)")
```


</details>

### c. Explain why the following matches the second 100 and not the first.

<details>
<summary class="answer">Click for answer</summary>
*answer:* It looks for one or more digits not followed by either a digit or space followed by `dollars`


```{r}
str_view(string1, "\\d+(?!\\d| dollars)")
```


For parts d and e, please take a look at `string2`. 


```{r}
string2 <- "USD100 PESO100"
```


</details>

### d. Explain why the following matches the first 100 and not the second.

<details>
<summary class="answer">Click for answer</summary>
*answer:* It looks for exactly 3 digits preceded by USD


```{r}
str_view(string2, "(?<=USD)\\d{3}")
```

</details>

### e. Explain why the following matches the second 100 and not the first.

<details>
<summary class="answer">Click for answer</summary>
*answer:* It looks for exactly 3 digits that is not preceded by USD


```{r}
str_view(string2, "(?<!USD)\\d{3}")
```

</details>

-------------------------------------------------------------------

##  Group Activity 3

```{r}
tweets<- read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/TrumpTweetData.csv")
```

### a. What proportion of tweets (text) mention "America"?

```{r}
tweets %>% 
  summarize(prop = mean(str_detect(str_to_title(text), "America"))) 
```

<details>
<summary class="answer">Click for answer</summary>
*Answer:* About 10% of tweets mention "America".

</details>

### b. What proportion of these tweets include "great"?

```{r}
tweets %>% filter(str_detect(str_to_title(text), "America")) %>%
  summarize(prop = mean(str_detect(str_to_lower(text), "great"))) 
```

<details>
<summary class="answer">Click for answer</summary>
*Answer:* About 40% of tweets mention "great".

</details>

### c. What proportion of the tweets mention `@`?

```{r}
tweets %>% mutate(ct = str_count(text, "@")) %>%
  select(text, ct) %>%
  summarize(prop = mean(ct>0))
```

<details>
<summary class="answer">Click for answer</summary>
*Answer:* About 32% of tweets mention `@`.

</details>

### d. Remove the tweets having mentions `@`.

```{r}
Mentions <- c("@[^\\s]+")

tw_noMentions <- tweets %>% mutate(textNoMentions = str_replace_all(text, Mentions, ""))
tw_noMentions$text[38]
tw_noMentions$textNoMentions[38]
```

<details>
<summary class="answer">Click for answer</summary>
*Answer:* `@:` This part of the pattern matches the "@" symbol, which usually indicates the beginning of a mention in a tweet.
[^\s]+: This part of the pattern matches one or more characters that are NOT whitespaces. The `^` inside the square brackets `[ ]` negates the character class (meaning it matches any character that is NOT in the specified class). The double backslash `\\` is used to escape the backslash in the R string, so the pattern `\\s` represents the whitespace character class `\s`. Finally, the + indicates that the pattern should match one or more occurrences of the non-whitespace characters. Together, this regular expression pattern `@[^\\s]+` matches any mention in a tweet, which usually starts with "@" followed by one or more non-whitespace characters.

</details>

### e. What poportion of tweets originated from an iPhone?


```{r}
tweets %>% group_by(source) %>% summarize(count = n()) %>%
  mutate(prop = count / sum(count)) %>%  filter(source == "iPhone")
```


<details>
<summary class="answer">Click for answer</summary>
*Answer:* About 42% of the tweets originated from an iPhone.

</details>
