# Class Activity 7

```{r}
# load the necessary libraries
library(tidyverse)
library(babynames)
```

## Problem 1: Boolean Operators

Use Boolean operators to alter the code below to return only the rows that contain:

### a. Girls named Rhea

<details>
<summary class="answer">Click for answer</summary>

```{r}
filter(babynames, name == "Rhea", sex == "F")
babynames %>%  filter(name == "Rhea", sex == "F")
```

</details>

### b. Names that were used by exactly 5 or 6 children in 1990

<details>
<summary class="answer">Click for answer</summary>

```{r}
filter(babynames, year == 1990, n == 5 | n == 6)
babynames %>% filter(year == "1990", n == 5 | n == 6)
```

</details>

### c. Names that are one of Apple, Yoroi, Ada

<details>
<summary class="answer">Click for answer</summary>

```{r}
filter(babynames, name == "Apple" | name == "Yoroi" | name == "Ada")
```

</details>

### d. Store the data tibble in part c into a new tibble and change all the character columns to upper case. Also, rename the `n` variable to `count`.

<details>
<summary class="answer">Click for answer</summary>

```{r}
aya <- babynames %>% filter(name == "Apple" | name == "Yoroi" | name == "Ada")
aya %>% mutate_if(is.character, toupper)
aya %>% mutate_at(vars(name), toupper)
aya %>% rename(count = n)
```

</details>

### e. Change all the column names to upper case in the previous problem.

<details>
<summary class="answer">Click for answer</summary>

```{r}
aya %>% rename_at(vars(year:prop), toupper)
```

</details>

### f. What do these commands do?

```{r}
polluted_cities <- tribble(
       ~city,   ~size, ~amount, 
  "New York", "large",      23,
  "New York", "small",      14,
    "London", "large",      22,
    "London", "small",      16,
   "Beijing", "large",      121,
   "Beijing", "small",      56
)

polluted_cities
```



```r
polluted_cities %>% select_if(is.numeric) #1
polluted_cities %>% rename_all(toupper) #2
polluted_cities %>% rename_if(is.character, toupper) #3
polluted_cities %>% rename_at(vars(contains("it")), toupper) #4
```

<details>
<summary class="answer">Click for answer</summary>

*answer:*

1. Selects all numeric columns from the polluted_cities dataset.
2. Renames all column names in the polluted_cities dataset to uppercase.
3. Renames column names with character data type in the polluted_cities dataset to uppercase.
4. Renames column names containing "it" in the polluted_cities dataset to uppercase.

```{r}
polluted_cities %>% select_if(is.numeric) #1
polluted_cities %>% rename_all(toupper) #2
polluted_cities %>% rename_if(is.character, toupper) #3
polluted_cities %>% rename_at(vars(contains("it")), toupper) #4
```

</details>

-----------------------------------------------------------------------------------

Let's look at an interesting example on how to join related information on various artists, bands, songs, and their labels.

```{r}
artists <- tibble(first = c("Jimmy", "George", "Mick", "Tom", "Davy", "John",
                            "Paul", "Jimmy", "Joe", "Elvis", "Keith", "Paul", 
                            "Ringo", "Joe", "Brian", "Nancy"), 
                  last = c("Buffett", "Harrison", "Jagger", "Jones", "Jones", 
                           "Lennon", "McCartney", "Page", "Perry", "Presley",
                           "Richards", "Simon", "Starr", "Walsh", "Wilson", "Wilson"), 
                  instrument = c("Guitar", "Guitar", "Vocals", "Vocals", "Vocals",
                                 "Guitar", "Bass", "Guitar", "Guitar", "Vocals", "Guitar", 
                                 "Guitar", "Drums", "Guitar", "Vocals", "Vocals"))


bands <- tibble(first = c("John", "John Paul", "Jimmy", "Robert", "George", "John", 
                          "Paul", "Ringo", "Jimmy", "Mick", "Keith", "Charlie", "Ronnie"), 
                last = c("Bonham", "Jones", "Page", "Plant", "Harrison", "Lennon",
                         "McCartney", "Starr", "Buffett", "Jagger", "Richards", "Watts", "Wood"), 
                band = c("Led Zeppelin", "Led Zeppelin", "Led Zeppelin", "Led Zeppelin",
                         "The Beatles", "The Beatles", "The Beatles", "The Beatles",
                         "The Coral Reefers", "The Rolling Stones", "The Rolling Stones",
                         "The Rolling Stones", "The Rolling Stones"))

albums <- tibble(album = c("A Hard Day's Night", "Magical Mystery Tour", "Beggar's Banquet",
                           "Abbey Road", "Led Zeppelin IV", "The Dark Side of the Moon", "Aerosmith",
                           "Rumours", "Hotel California"),
                 band = c("The Beatles", "The Beatles", "The Rolling Stones", "The Beatles", 
                          "Led Zeppelin", "Pink Floyd", "Aerosmith", "Fleetwood Mac", "Eagles"), 
                 year = c(1964,1967,1968,1969,1971,1973,1973,1977,1982))


songs <- tibble(song = c("Come Together", "Dream On", "Hello, Goodbye", "It's Not Unusual"),
                album  = c("Abbey Road", "Aerosmith", "Magical Mystery Tour", "Along Came Jones"), 
                first = c("John", "Steven", "Paul", "Tom"), 
                last = c("Lennon", "Tyler", "McCartney", "Jones"))


labels <- tibble(album = c("Abbey Road", "A Hard Days Night", "Magical Mystery Tour",
                           "Led Zeppelin IV", "The Dark Side of the Moon", "Hotel California", 
                           "Rumours", "Aerosmith", "Beggar's Banquet"), 
                 label = c("Apple", "Parlophone", "Parlophone", "Atlantic", "Harvest",
                           "Asylum", "Warner Brothers", "Columbia", "Decca"))

```



Let's take a glimpse of the tibbles `artists` and `bands`. Notice that there are different number of rows in the dataset.

```{r}
glimpse(artists)
glimpse(bands)
glimpse(albums)
glimpse(songs)
glimpse(labels)
```


## Problem 2: Joining artists and bands data


### a. Join the artists and bands tibbles using `left_join()`, `right_join()`, and `full_join()`. Verify that the datasets obtained from `left_join()` and `right_join()` are the same using `setequal()`.

<details>
<summary class="answer">Click for answer</summary>

```{r}
bands2 <- left_join(bands, artists, by = c("first", "last"))
bands3 <- right_join(artists, bands, by = c("first", "last"))
full_bands <- full_join(artists, bands, by = c("first", "last"))

# Check if the datasets are the same
setequal(bands2, bands3)
```

</details>

### b. Use the pipe operator, `%>%`, to create one table that combines all information from `artists`, `bands`, `albums`, `songs`, and `labels.`

<details>
<summary class="answer">Click for answer</summary>

```{r}
all_info <- artists %>% 
  full_join(bands, by = c("first", "last")) %>% 
  full_join(songs, by = c("first", "last")) %>% 
  full_join(albums, by = c("album", "band")) %>%
  full_join(labels, by = c("album"))
all_info
```

</details>

-----------------------------------------------------------------------------------

## Problem 3: Filtering and counting rows in the data

### a. Collect artists that have songs provided, and return rows of artists that don't have bands info.

<details>
<summary class="answer">Click for answer</summary>

```{r}
# Artists with songs
artists_with_songs <- artists %>% 
  semi_join(songs, by = c("first", "last"))

# Artists without bands info
artists_without_bands <- artists %>% 
  anti_join(bands, by = c("first","last"))

artists_with_songs
artists_without_bands
```

</details>

### b. Collect the albums made by a band, count the number of rows, find the rows of songs that match a row in labels, and count the number of rows.

<details>
<summary class="answer">Click for answer</summary>

```{r}
# Albums made by a band
albums_by_band <- albums %>% semi_join(bands, by = "band")
n_albums_by_band <- nrow(albums_by_band)

# Rows of songs that match a row in labels
songs_with_labels <- songs %>% semi_join(labels, by = "album")
n_songs_with_labels <- nrow(songs_with_labels)

n_albums_by_band
n_songs_with_labels
```

</details>

