# Class Activity 9

```{r}
# load the necessary libraries
library(tidyverse)
```


## Your Turn 1

### a) `read_csv()`

Use `read_csv()` to import the `desserts` data set from https://raw.githubusercontent.com/deepbas/statdatasets/main/desserts.csv. Use `glimpse` to see if the data import is alright.

```{r}
url <- "https://raw.githubusercontent.com/deepbas/statdatasets/main/desserts.csv"
desserts <- read_csv(url)
glimpse(desserts)
```

### b) Are there any issues with the data import? If so, what are they?

<details>
<summary class="answer">Click for answer</summary>

*Answer:* Based on the output of glimpse, we can see that the 'technical' column should be a numeric column and the 'uk_airdate' column should be a date column. We can also identify any issues with missing values.


```{r}
# your r-code

desserts <- read_csv(url,
  col_types = list(
    technical = col_number(),
    uk_airdate = col_date()
  )
)

problems(desserts)
```

</details>


### c) Import the dataset with correct data types, if needed. Fix the problems, if any.

<details>
<summary class="answer">Click for answer</summary>

```{R}
desserts <- read_csv(url,
  col_types = list(
    technical = col_number(),
    uk_airdate = col_date()
  )
)

problems(desserts)
```



```{r}
desserts <- read_csv(url,
    col_types = list(
    technical = col_number(), 
    uk_airdate = col_date(format = "%d %B %Y")
  ) 
)

problems(desserts)
```



```{r}
desserts <- read_csv(url,
  col_types = list(
    technical = col_number(), 
    uk_airdate = col_date(format = "%d %B %Y")
  ),
  na = c("", "NA", "N/A")
)

problems(desserts)
```

</details>


## Your Turn 2

Use the appropriate `read_<type>()` function to import the following data sets:

- `https://deepbas.io/data/simple-1.dat`

- `https://deepbas.io/data/mild-1.csv`

- `https://deepbas.io/data/tricky-1.csv`

- `https://deepbas.io/data/tricky-2.csv`


Identify and fix any issues you encounter.

### a)  Importing simple data:

<details>
<summary class="answer">Click for answer</summary>

```{r}
simple1 <- readr::read_csv("https://deepbas.io/data/simple-1.dat")
problems(simple1)
```

</details>

### b) Importing mildly tricky data:

<details>
<summary class="answer">Click for answer</summary>

```{r}
mild1 <- readr::read_delim("https://deepbas.io/data/mild-1.csv", delim = "|")
problems(mild1)
```

</details>

### c) Importing tricky data 1:

<details>
<summary class="answer">Click for answer</summary>

```{r}
tricky1 <-  read_csv("https://deepbas.io/data/tricky-1.csv")
problems(tricky1)

# Fix missing values
tricky1[3, ] <- c(tricky1[3, 1:2], NA, tricky1[3, 3:4])
tricky1[6, ] <- c(tricky1[4, 1], NA, tricky1[4, 3:5])
```

</details>


### d) Importing tricky data 2:

<details>
<summary class="answer">Click for answer</summary>

```{r}
tricky2 <- read_csv("https://deepbas.io/data/tricky-2.csv")
problems(tricky2)

# Fix missing values
tricky2_part1 <- read_csv("https://deepbas.io/data/tricky-2.csv", n_max = 7) %>%
  separate(city, c("city", "state"), sep = ",") %>%
  select(-c(7))

tricky2_part2 <- read_csv(
  "https://deepbas.io/data/tricky-2.csv",
  skip = 8, 
  col_names = c("iata", "airport", "city", "state",  "latitude", "longitude")
)

# Combine parts
data_combined <- full_join(tricky2_part1, tricky2_part2)

```

</details>
