# Class Activity 20

```{r}
# load the necessary libraries
library(tidyverse)
library(shiny)
library(rvest)
library(polite)
library(leaflet)
library(sp)
library(maptools)
library(rgeos)
library(stringr)
library(RColorBrewer)
library(terra)
library(raster)
```


## Group Activity 1

### a. Explore COVID-19 vaccination rates across the Midwest with this R script, which scrapes data, processes it, and creates an interactive, state-specific leaflet map for clear visualization.


<details>
<summary class="answer">Click for answer</summary>
*Answer:* 

```{r, eval=FALSE}
# Scrape the data
covid_final <- read_html("https://usafacts.org/visualizations/covid-vaccine-tracker-states/state/minnesota") %>%
  html_elements(css = "table") %>% html_table() %>% .[[1]] %>%
  janitor::clean_names() %>%
  mutate_at(2:4, parse_number) %>% mutate(state = str_to_lower(state)) %>%
  filter(state %in% c("ohio", "indiana", "michigan", "illinois", "missouri", "wisconsin", "minnesota", 
                                     "iowa", "kansas", "nebraska", "south dakota", "north dakota"))

# Midwest
USA_Midwest <- maps::map("state", 
                         regions = c("ohio", "indiana", "michigan", "illinois", "missouri", "wisconsin", "minnesota", 
                                     "iowa", "kansas", "nebraska", "south dakota", "north dakota"), 
                         plot = FALSE, fill = TRUE) %>%
  map2SpatialPolygons(IDs = str_remove(.$names, "(?=:).+"))

# Merge the data and the map
map <- SpatialPolygonsDataFrame(USA_Midwest, covid_final, match.ID = FALSE)

# Create bins and color palette
bins <- seq(min(map$percent_fully_vaccinated), max(map$percent_fully_vaccinated), length.out = 6)
pal <- colorBin("viridis", domain = map$percent_fully_vaccinated, bins = bins)

# Create labels
labels <- sprintf("<strong> %s </strong> <br/> Observed: %s", str_to_upper(map$state), map$percent_fully_vaccinated) %>%
  lapply(htmltools::HTML)

# Plot the map
leaflet(map) %>% 
  addTiles() %>% 
  setView(lng = -93.1616, lat = 44.4583, zoom = 4) %>%
  addPolygons(
    color = "grey", 
    weight = 1,
    fillColor = ~pal(percent_fully_vaccinated), 
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(weight = 5),
    label = labels
  ) %>%
  addLegend(
    pal = pal, 
    values = ~percent_fully_vaccinated, 
    opacity = 0.5, 
    title = "Percent Vaccn.", 
    position = "bottomright"
  )

```

</details>


----------------------------------------------------------------------------------------------

## Group Activity 2

### a. Explore the COVID-19 vaccination status across various US regions using this interactive Shiny app. The script extracts data from USAFacts, presents user options to customize the map projection, color palette, and region, and dynamically generates a Leaflet map to visualize the selected data.



<details>
<summary class="answer">Click for answer</summary>
*Answer:* 

```{r, eval=FALSE}
# Scrape the data
covid_final <- read_html("https://usafacts.org/visualizations/covid-vaccine-tracker-states/state/minnesota") %>%
  html_elements(css = "table") %>% html_table() %>% .[[1]] %>%
  janitor::clean_names() %>%
  mutate_at(2:4, parse_number) %>% mutate(state = str_to_lower(state))



# UI
ui <- fluidPage(
  titlePanel("Map of States and their vaccination status"),
  sidebarLayout(
    sidebarPanel(
      selectInput("myvar", "What to project?", choices = names(covid_final)[2:4], selected = names(covid_final)[2]),
      selectInput("palette", "Choose color palette:", choices = c("viridis", "magma", "inferno", "plasma", "cividis")),
      selectInput("region", "Select region:", choices = c("Northeast", "Southeast", "Midwest", "South", "West")),
      actionButton("goButton", "Update map")
    ),
    mainPanel(
      leafletOutput("myplot")
    )
  ) 
)


# UI
ui <- fluidPage(
  titlePanel("Map of States and their vaccination status"),
  sidebarLayout(
    sidebarPanel(
      selectInput("myvar", "What to project?", choices = names(covid_final)[2:4], selected = names(covid_final)[2]),
      selectInput("palette", "Choose color palette:", choices = c("viridis", "magma", "inferno", "plasma", "cividis")),
      selectInput("region", "Select region:", choices = c("Northeast", "Southeast", "Midwest", "South", "West")),
      actionButton("goButton", "Update map")
    ),
    mainPanel(
      leafletOutput("myplot")
    )
  ) 
)

# Server
server <- function(input, output, session) {
  map_data <- eventReactive(input$goButton, {
    req(input$region)
    
    region_states <- switch(input$region,
                        "Northeast" = c("maine", "new hampshire", "vermont", "massachusetts", "rhode island", 
                                        "connecticut", "new york", "pennsylvania", "new jersey"),
                        "Southeast" = c("delaware", "maryland", "virginia", "west virginia", "north carolina", 
                                        "south carolina", "georgia", "florida", "kentucky", "tennessee", 
                                        "mississippi", "alabama", "arkansas", "louisiana"),
                        "Midwest" = c("ohio", "indiana", "michigan", "illinois", "wisconsin", "minnesota", 
                                      "iowa", "missouri", "kansas", "nebraska", "south dakota", "north dakota"),
                        "South" = c("oklahoma", "texas", "new mexico", "arizona"),
                        "West" = c("colorado", "wyoming", "montana", "idaho", "utah", "nevada", "california", 
                                   "oregon", "washington"))
                       

    
    covid_region <- isolate(covid_final %>%
      filter(state %in% region_states))
    
    USA_region <- maps::map("state", regions = region_states, plot = FALSE, fill = TRUE) %>%
      map2SpatialPolygons(IDs = str_remove(.$names, "(?=:).+"))
    
    SpatialPolygonsDataFrame(USA_region, covid_region, match.ID = FALSE)
  }, ignoreNULL = FALSE)
  
  output$myplot <- renderLeaflet({
    req(map_data()) # ensure that map_data() is available
    
    bins <- seq(min(map_data()[[input$myvar]]), max(map_data()[[input$myvar]]), length.out = 6)
    pal <- colorBin(input$palette, domain = map_data()[[input$myvar]], bins = bins)
    
    labels <- sprintf("<strong> %s </strong> <br/> Observed <strong> %s </strong>  : %s", str_to_upper(map_data()$state), input$myvar, map_data()[[input$myvar]]) %>% 
      lapply(htmltools::HTML)
    
    m <- leaflet(map_data()) %>%
      addTiles() %>%  setView(lng = -93.1616, lat = 40.4583, zoom = 3.5) %>%
      addPolygons(
        color = "grey", 
        weight = 1,
        fillColor = ~pal(get(input$myvar)), 
        fillOpacity = 0.7,
        highlightOptions = highlightOptions(weight = 5),
        label = labels
      ) %>%
      addLegend(
        pal = pal, 
        values = ~get(input$myvar), 
        opacity = 0.5, 
        title = str_to_title(input$myvar), 
        position = "bottomright"
      )
    m
  })
}

# Run the app
shinyApp(ui = ui, server = server)
```

</details>


### b. Modify the app in 2a to add an "All" option in the "Select region" input to allow users to view all states at once. 

```{r}
# your r-code


```


### c. Open an account on [shinyapps.io](https://www.shinyapps.io/admin/#/dashboard) and follow the steps to deploy one of the apps to [shinyapps.io](https://www.shinyapps.io/admin/#/dashboard).


