# Class Activity 22

```{r}
# load the necessary libraries
library(tidyverse) 
library(tidymodels)
library(mlbench)     # for PimaIndiansDiabetes2 dataset
library(janitor)
library(parsnip)
library(kknn)
library(ggthemes)
library(purrr)
library(forcats)
```



## Group Activity 1

Load the `mlbench` package to get `PimaIndiansDiabetes2` dataset.

```{r}
# Load the data - diabetes
data(PimaIndiansDiabetes2)
db <- PimaIndiansDiabetes2
db <- db %>% drop_na() %>% mutate(diabetes = fct_rev(factor(diabetes))) 
db_raw <- db %>% select(glucose, insulin, diabetes)

```


a. Split the data `75-25` into training and test set using the following code. 

<details>
<summary class="answer">Click for answer</summary>
*Answer:* 

```{r}
set.seed(123)

db_split <- initial_split(db, prop = 0.75)

# Create training data
db_train <- db_split %>% training()

# Create testing data
db_test <- db_split %>%  testing()
```
  
</details>

  
b. Follow the steps to train a 7-NN classifier using the `tidymodels` toolkit 
    
<details>
<summary class="answer">Click for answer</summary>
*Answer:* 

```{r}
# define recipe and preprocess the data
db_recipe <- recipe(diabetes ~ ., data = db_raw) %>%
  step_scale(all_predictors()) %>%
  step_center(all_predictors()) %>%
  prep()
```
    

```{r}
# specify the model
db_knn_spec7 <- nearest_neighbor(mode = "classification",
                             engine = "kknn",
                             weight_func = "rectangular",
                             neighbors = 7)
```


```{r}
# define the workflow
db_workflow <- workflow() %>% 
  add_recipe(db_recipe) %>%
  add_model(db_knn_spec7)
```


```{r}
# fit the model
db_fit <- fit(db_workflow, data = db_train)
```
    
</details>

c. Classify the penguins in the `test` data frame. 


<details>
<summary class="answer">Click for answer</summary>
*Answer:* 

```{r}
test_features <- db_test %>% select(glucose, insulin) 
db_pred <- predict(db_fit, test_features)

db_results <- db_test %>% 
  select(glucose, insulin, diabetes) %>% 
  bind_cols(predicted = db_pred)

head(db_results, 6)
```

</details>


-----------------------------------------------------------------------------------------------

## Group Activity 2

Calculate the accuracy, sensitivity, specificity, and positive predictive value by hand using the following confusion matrix.

```{r}
conf_mat(db_results, truth = diabetes, estimate = .pred_class)
```

<details>
<summary class="answer">Click for answer</summary>
*Answer:* 

```{r}
accuracy(db_results, truth = diabetes,
         estimate = .pred_class)

sens(db_results, truth = diabetes,
         estimate = .pred_class)

spec(db_results, truth = diabetes,
         estimate = .pred_class)

ppv(db_results, truth = diabetes,
         estimate = .pred_class)
```

</details>


## Extra: Code to recreate the plot in the `slides` for the `diabetes` dataset.


<details>
<summary class="answer">Click for answer</summary>
*Answer:* 


```{r}
metrics_for_k <- function(k, db_train, db_test){
db_knn_spec <- nearest_neighbor(mode = "classification",
                             engine = "kknn",
                             weight_func = "rectangular",
                             neighbors = k)

db_knn_wkflow <- workflow() %>%
  add_recipe(db_recipe) %>%
  add_model(db_knn_spec)

db_knn_fit <- fit(db_knn_wkflow, data = db_train)
test_features <- db_test %>% select(glucose, insulin)
nn1_pred <- predict(db_knn_fit, test_features, type = "raw")

db_results <- db_test %>% 
  select(diabetes) %>% 
  bind_cols(predicted = nn1_pred)
custom_metrics <- metric_set(accuracy, sens, spec, ppv)

metrics <- custom_metrics(db_results,
               truth = diabetes,
               estimate = predicted) 
metrics <- metrics %>% select(-.estimator) %>% mutate(k = rep(k,4))

return(list = metrics)
}
```

<!--

```{r}
k <- seq(1,40, by=1)
optim.results <- purrr::map_dfr(k, ~metrics_for_k(.x, db_train, db_test)) 
```

-->

```{r}
optim.results %>%
  ggplot(aes(x = k, y = .estimate, color = forcats::fct_reorder2(.metric, k, .estimate ))) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  theme_minimal() +
  ggthemes::scale_color_wsj() + 
  scale_x_continuous(breaks = k) +
  theme(panel.grid.minor.x = element_blank(),
        axis.text=element_text(size=6, angle = 20))+
  labs(color='Metric', y = "Estimate", x = "K") 
```

</details>
