# Class Activity 18

```{r}
# load the necessary libraries
library(tidyverse)
library(shiny)
library(readr)
library(janitor)
library(purrr)
library(lubridate)
library(DT)
library(ggthemes)
library(rvest)
library(polite)
```



## Introduction to Shiny Apps

Shiny is a powerful R package that enables the creation of interactive web applications directly from R. A Shiny app consists of two main components:


- *User Interface (UI):* Defines the layout and appearance of your app.

- *Server Function:* Contains the logic to manage user inputs and generate outputs.


#### Step 1: Setting Up a Basic Shiny App

Start with a simple "Hello World!" app to understand the basic structure.


```{r, eval=FALSE}
ui <- fluidPage("Hello World!")
server <- function(input, output) {}
shinyApp(ui = ui, server = server, options = list(height = 200))
```


<br>

#### Step 2: Enhancing the UI with More Information

Add more elements to the UI to make it informative.

```{r, eval=FALSE}
ui <- fluidPage(
  titlePanel("Tracking Covid in Minnesota"),
  h1("A Detailed Overview"),
  p("elements1", "elements2"),
  br(),
  p("Detailed Information:", strong("Critical Data Points"))
)

server <- function(input, output) {}
shinyApp(ui = ui, server = server, options = list(height = 200))
```

<br>

#### Step 3: Structuring the App with a Sidebar Layout

Organize the app's layout using sidebarLayout to separate inputs from results.

```{r, eval=FALSE}
ui <- fluidPage(
  titlePanel("Tracking Covid in Minnesota"),
  sidebarLayout(
    sidebarPanel(p("Inputs will be placed here")),
    mainPanel(p("Results will be displayed here"))
  )
)

server <- function(input, output) {}
shinyApp(ui = ui, server = server, options = list(height = 200))
```

#### Step 4: Reading and Preparing Data

Fetch and prepare Covid data for Minnesota for analysis.

```{r, eval=FALSE}
table_usafacts <- bow(url = "https://usafacts.org/visualizations/coronavirus-covid-19-spread-map/state/minnesota") %>%
  scrape() %>%  html_nodes("a") %>%       # find all links
  html_attr("href") %>%     # get the url
  str_subset(".csv")        # find those that end in csv

covid_data <- read_csv(table_usafacts[2]) %>%
  filter(State == "MN") %>%
  select(-countyFIPS, -StateFIPS, -State) %>%
  pivot_longer(-1, names_to = "Dates", values_to = "Cases") %>%
  clean_names() %>%
  mutate(dates = ymd(dates),
         counties = factor(str_remove(county_name, " County"))) %>%
  select(-county_name)

head(covid_data, 10) %>% knitr::kable()
```


#### Step 5: Incorporating User Inputs

Enhance the UI with inputs for filtering data based on user selection.

```{r, eval=FALSE}
ui <- fluidPage(
  titlePanel("Tracking Covid in Minnesota"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("monthInput", "Select Month Range", min = 1, max = 12, value = c(3, 6)),
      selectInput("countyInput", "Select County", choices = unique(covid_data$counties))
    ),
    mainPanel(
      plotOutput("covidPlot"),
      tableOutput("covidTable")
    )
  )
)


server <- function(input, output) {}

shinyApp(ui = ui, server = server)

```

#### Step 6: Developing the Server Logic
 
Implement server logic to respond to user inputs and generate outputs.

```{r, eval=FALSE}
server <- function(input, output) {
  # Filter data based on inputs
  filtered_data <- reactive({
    covid_data %>%
      filter(counties == input$countyInput, 
             month(dates) >= input$monthInput[1], 
             month(dates) <= input$monthInput[2])
  })
  
  # Generate plot
  output$covidPlot <- renderPlot({
    ggplot(filtered_data(), aes(x = dates, y = cases, color = counties)) +
      geom_line() +
      theme_minimal() +
      labs(title = "Covid Cases Over Time", y = "Number of Cases")
  })
  
  # Generate table
  output$covidTable <- renderTable({
    filtered_data()
  })
}

# Launch the app
shinyApp(ui = ui, server = server)
```



### Implementation 2

```{r, eval=FALSE}
ui1 <- fluidPage(
    titlePanel("Tracking Covid in Minnesota"),
    sidebarLayout(
        sidebarPanel(
          sliderInput("monthInput", "Month", 0, 12, c(3, 6)),
          radioButtons("yearInput", "Year",
                  choices = c("2020", "2021", "2022", "2023"),
                  selected = "2020"),
          selectInput(inputId = "dv", label = "County",
                        choices = levels(covid_data$counties),
                        selected = c("Aitkin"))
    ),
        mainPanel(
            plotOutput(outputId = "plot"), br(),
            DT::dataTableOutput(outputId = "table")
        )
    )
)

```



```{r, eval=FALSE}
server1 <- function(input, output) {
    filtered_data <- reactive({
        subset(covid_data,
               counties %in% input$dv &
               month >= input$monthInput[1] & month <= input$monthInput[2] & 
          year == input$yearInput) })

    output$plot <- renderPlot({
        ggplot(filtered_data(), aes(x=dates, y=cases, color=counties)) + theme_economist_white()+
                geom_point(alpha=0.5, color = "blue") + theme(legend.position = "none") +
                    ylab("Number of Cases") + xlab("Date")})

    output$table <- DT::renderDataTable({
        filtered_data()})
        
}
```



```{r, out.width = "300%", , eval=FALSE}
app1 <- shinyApp(ui = ui1, server = server1, options = list(height = 1200))
app1
```
