# R Markdown Basics

An R Markdown file (.Rmd file) combines R commands and written analyses, which are 'knit' together into an HTML, PDF, or Microsoft Word document. 

An R Markdown file contains three essential elements:

- Header: The header (top) of the file contains information like the document title, author, date and your preferred output format (pdf_document, word_document, or html_document).

- Written analysis: You write up your analysis after the header and embed R code where needed. The online help below shows ways to add formatting details like bold words, lists, section labels, etc to your final pdf/word/html document. For example, adding ** before and after a word will bold that word in your compiled document. 

- R chunks: R chunks contain the R commands that you want evaluated. You embed these chunks within your written analysis and they are evaluated when you compile the document.


## Some R Markdown helpers

Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

### Lists in R Markdown:

You can use asterisk mark to provide emphasis, such as `*italics* or **bold**`. You can create lists with a dash:

```r
- Item 1
- Item 2
- Item 3
  + Subitem 1
* Item 4
```
to produce

- Item 1
- Item 2
- Item 3
  + Subitem 1
* Item 4


You can embed Latex equations in-line, `$\frac{1}{n} \sum_{i=1}^{n} x_{i}$` to produce $\frac{1}{n} \sum_{i=1}^{n} x_{i}$  or in a new line as `$$\text{Var}(X) = \frac{1}{n-1}\sum_{i-1}^{n} (x_{i} - \bar{x})^2$$` to produce $$\text{Var}(X) = \frac{1}{n-1}\sum_{i-1}^{n} (x_{i} - \bar{x})^2$$



### Embed an R code chunk:

Use the following

````
```r`r ''`
Use back ticks to 
create a block of code
```
````

to produce:

```
Use back ticks to 
create a block of code
```

You can also evaluate and display the results of R code. Each tasks can be accomplished in a suitably labeled chunk like the following:


```{r cars}
summary(cars)
fit <- lm(dist ~ speed, data = cars)
fit
```


### Including Plots:

You can also embed plots. See Figure \@ref(fig:pie) for example:

```{r pie, fig.cap='A fancy pie chart.', fig.height=7, fig.width=7}
par(mar = c(0, 1, 0, 1))
pie(
  c(280, 60, 20),
  c('Sky', 'Sunny side of pyramid', 'Shady side of pyramid'),
  col = c('#0292D8', '#F7EA39', '#C4B632'),
  init.angle = -50, border = NA
)
```

(Credit: Yihui Xie)

### Read in data files:

```{r read_data}
simple_data <- read.csv("https://deepbas.io/data/simple-1.dat", )
summary(simple_data) 
```


```{r pretty_table}
knitr::kable(simple_data)
```

### Hide the code:

If we enter the `echo = FALSE` option in the R chunk (see the .Rmd file). This prevents the R code from being printed to your document; you just see the results.


```{r pretty_table2, echo=FALSE}
knitr::kable(simple_data)
```



